Equivalence Testing and the Second Generation \emph{P}-Value

### 2019-03-09 manuscript.R1.tex
R1

\documentclass[,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Equivalence Testing and the Second Generation P-Value},
            pdfauthor={Daniël Lakens~\& Marie Delacre},
            pdfkeywords={equivalence testing, second generation \emph{p}-values, hypothesis
testing, TOST, statistical inference},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\shorttitle{TOST vs. SGPV}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Eindhoven University of Technology, Eindhoven, The Netherlands\\\textsuperscript{2} Service of Analysis of the Data, Université Libre de Bruxelles, Belgium}
\keywords{equivalence testing, second generation *p*-values, hypothesis testing, TOST, statistical inference\newline\indent Word count: }
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\abstract{
To move beyond the limitations of null-hypothesis tests, statistical
approaches have been developed where the observed data is compared
against a range of values that are equivalent to the absence of a
meaningful effect. Specifying a range of values around zero allows
researchers to statistically reject the presence of effects large enough
to matter, and prevents practically insignificant effects from being
interpreted as a statistically significant difference. We compare the
behavior of the recently proposed second generation \emph{p}-value
(Blume, D'Agostino McGowan, Dupont, \& Greevy, 2018) with the more
established Two One-Sided Tests (TOST) equivalence testing procedure
(Schuirmann, 1987). We show that the two approaches yield almost
identical results under optimal conditions. Under suboptimal conditions
(e.g., when the confidence interval is wider than the equivalence range,
or when confidence intervals are asymmetric) the second generation
\emph{p}-value becomes difficult to interpret as a descriptive
statistic. The second generation \emph{p}-value is interpretable in a
dichotomous manner (i.e., when the SGPV equals 0 or 1 because the
confidence intervals lies completely within or outside of the
equivalence range), but this dichotomous interpretation does not require
calculations. We conclude that equivalence tests yield more consistent
\emph{p}-values, distinguish between datasets that yield the same second
generation \emph{p}-value, and allow for easier control of Type I and
Type II error rates.

@@

To test predictions researchers predominantly rely on null-hypothesis
tests. This statistical approach can be used to examine whether observed
data is sufficiently surprising under the null hypothesis to reject an
effect that equals exactly zero. Null-hypothesis tests have an important
limitation, in that this procedure can only reject the hypothesis that
there is no effect, while scientists should also be able to provide
statistical support for \emph{equivalence}. When testing for equivalence
researchers aim to examine whether an observed effect is too small to be
considered meaningful, and therefore is practically equivalent to zero.
By specifying a range around the null hypothesis of values that are
deemed practically equivalent to the absence of an effect (i.e., 0 ±
0.3) the observed data can be compared against an \emph{equivalence
range} and researchers can test if a meaningful effect is absent (Hauck
\& Anderson, 1984; Kruschke, 2018; Rogers, Howard, \& Vessey, 1993;
Serlin \& Lapsley, 1985; Spiegelhalter, Freedman, \& Parmar, 1994;
Wellek, 2010; Westlake, 1972).

Second generation \emph{p}-values (SGPV) were recently proposed to as a
descriptive statistic that represents \enquote{the proportion of
data-supported hypotheses that are also null hypotheses} (Blume et al.,
2018). The researcher specifies an equivalence range around a null
hypothesis of values that are considered practically equivalent to the
null hypothesis. The SGPV measures the degree to which a set of
data-supported parameter values falls within the interval null
hypothesis. If the estimation interval falls completely within the
equivalence range, the SGPV is 1. If the confidence interval falls
completely outside of the equivalence range, the SGPV is 0. Otherwise
the SGPV is a value between 0 and 1 that expresses the overlap of
data-supported hypotheses and the equivalence range. When calculating
the SGPV the set of data-supported parameter values can be represented
by a confidence interval (CI), although one could also choose to use
credible intervals or Likelihood support intervals (SI). When a
confidence interval is used, the SGPV and equivalence tests such as the
Two One-Sided Tests (TOST) procedure (Lakens, 2017; Meyners, 2012;
Quertemont, 2011; Schuirmann, 1987) appear to have close ties, because
both tests compare a confidence interval against an equivalence range.
Here, we aim to examine the similarities and differences between the
TOST procedure and the SGPV. We limit our analysis to normally
distributed continuous data.

The TOST procedure also relies on the confidence interval around the
effect. In the TOST procedure the data is tested against the lower
equivalence bound in the first one-sided test, and against the upper
equivalence bound in the second one-sided test (Lakens, Scheel, \&
Isager, 2018). For an excellent discussion of the strengths and
weaknesses of different frequentist equivalence tests, including
alternatives to the TOST procedure, see (Meyners, 2012). If both tests
statistically reject an effect as extreme or more extreme than the
equivalence bound, you can conclude the observed effect is practically
equivalent to zero from a Neyman-Pearson approach to statistical
inferences. Because one-sided tests are performed, one can also conclude
equivalence by checking whether the 1-2\(\times\)\(\alpha\) confidence
interval (e.g., when the alpha level is 0.05, a 90\% CI) falls
completely within the equivalence bounds. Because both equivalence tests
as the SGPV are based on whether and how much a confidence interval
overlaps with equivalence bounds, it seems worthwhile to compare the
behavior of the newly proposed SGPV to equivalence tests to examine the
unique contribution of the SGPV to the statistical toolbox.

@@

For example, in Figure \ref{fig:TOSTSGPV4} we have plotted four SGPV's.
From A to D the SGPV is 0.76, 0.81, 0.86, and 0.91. The difference in
the percentage of overlap between A and B (-0.05) is identical to the
difference in the percentage of overlap between C and D as the mean gets
0.1 closer to the test value (-0.05). As the observed mean in a
one-sample \emph{t}-test lies closer to the test value, from situation A
to D, the mean gets closer to the test value by 0.1) the difference in
the overlap changes uniformly. As we move the observed mean closer to
the test value in steps of 0.1 across A to D the \emph{p}-value
calculated for normally distributed data is not uniformly distributed.
The probability of observing data more extreme than the upper bound of 2
is (from A to D) 0.16, 0.12, 0.08, and 0.06. As we can see, the
difference between A and B (0.04) is not the same as the difference
between C and D (0.03). Indeed, the difference in \emph{p}-values is the
largest as you start at \emph{p} = 0.5 (when the observed mean falls on
the test value), which is why the line in Figure \ref{fig:TOSTSGPV1} is
the steepest at \emph{p} = 0.5. Note that where the SGPV reaches 1 or 0,
\emph{p}-values closely approximate 0 and 1, but never reach these
values.

@@

We can examine this situation by calculating the SGPV and performing the
TOST for a situation where sample sizes are small and the equivalence
range is narrow, such that the CI is more than twice as large as the
equivalence range (see Figure \ref{fig:TOSTSGPV6}). When the two
statistics are plotted against each other we can see where they are
unrelated (indicated by straight lines in the curve, see Figure
\ref{fig:TOSTSGPV7}). We see the SGPV is 0.5 for a range of observed
means where the \emph{p}-value from the equivalence test still varies.
It should be noted that in these calculations the \emph{p}-values for
the TOST procedure are \emph{never} smaller than 0.05 (i.e., they do not
get below 0.05 on the y-axis). In other words, we cannot conclude
equivalence based on any of the observed means. This happens because the
TOST \emph{p}-value is smaller than 0.05 only when the 90\% CI falls
completely between the upper and lower equivalence bounds. However, we
are examining a scenario where the 90\% CI is so wide that it never
falls completely within the two equivalence bounds, and thus the
equivalence test is never significant. As Lakens (2017) notes:
\enquote{in small samples (where CIs are wide), a study might have no
statistical power (i.e., the CI will always be so wide that it is
necessarily wider than the equivalence bounds).} None of the
\emph{p}-values based on the TOST procedure are below 0.05, and thus, in
the long run we have 0\% power.

@@

The \emph{p}-value from the TOST procedure and the SGPV are also
unrelated when the CI is wider than the equivalence range (so the
precision is low) and overlaps with the upper and lower equivalence
bound, but the CI is \emph{not} twice as wide as the equivalence range.
In the example below, we see that the CI is only 1.79 times as wide as
the equivalence bounds, but the CI overlaps with the lower and upper
equivalence bounds (Figure \ref{fig:TOSTSGPV8}). This means the SGPV is
not set to 0.5, but it is constant across a range of observed means.

@@

If we once more plot the two statistics against each other to see where
they are unrelated (indicated by straight lines in the curve), we see
the SGPV is 0.56 for a range of observed means where the \emph{p}-value
from the equivalence test still varies (Figure \ref{fig:TOSTSGPV10}).

@@

To conclude this section, there are three situations where the
\emph{p}-value from the TOST procedure is unrelated to the SGPV. In all
these situations the \emph{p}-value for the equivalence test
differentiates tests with different means, but the SGPV does not.
Therefore, as a purely continuous descriptive statistic, the SGPV is
more limited than the \emph{p}-value from the TOST procedure.

\subsection{The relation between equivalence tests and SGPV when for
asymmetrical confidence intervals around
correlations}\label{the-relation-between-equivalence-tests-and-sgpv-when-for-asymmetrical-confidence-intervals-around-correlations}

So far we have only looked at the relation between equivalence tests and
the SGPV when confidence intervals are symmetric (e.g., for confidence
intervals around mean differences). For correlations, which are bound
between -1 and 1, confidence intervals are only symmetric for a
correlation of exactly 0. The confidence interval for a correaltion
becomes increasingly asymmetric as the observed correlation nears -1 or
1. For example, with ten observations, an observed correlation of 0 has
a symmetric 95\% confidence interval ranging from -0.630 to 0.630, while
and observed correlation of 0.7 has an asymmetric 95\% confidence
interval ranging from 0.126 to 0.993. Note that calculating confidence
intervals for a correlation involves a Fisher's z-transformation, which
transforms values such that they are approximately normally distributed,
which allows one to compute symmetric confidence intervals, which are
then retransformed into a correlation, where the confidence intervals
are asymmetric if the correlation is not exactly zero.

@@

This example illustrates the difference between a proportion and a
probability. As long as data is normally distributed, there is always a
50\% probability of observing a correlation smaller or larger than the
true correlation, but the SGPV for this situation depends on how far
away the observed correlation is from 0. The further away from 0, the
larger the SGPV when the observed mean falls on the equivalence bound.
The SGPV is the proportion of values in a 95\% confidence interval that
overlap with the equivalence range, but not the probability that these
values will be observed. In the most extreme case (i.e., a sample size
of 4, and equivalence bounds set to r = -0.99 and 0.99, with an observed
correlation of 0.99) 58.10\% of the confidence interval overlaps with
the equivalence range, even though in the long run only 50\% of the
correlations observed in the future will fall in this range. It should
be noted that in larger sample sizes the SGPV is closer to 0.5 whenever
the observed correlation falls on the equivalence bound, but this
extreme example nevertheless clearly illustrates the difference between
question the SGPV answers, and the question a \emph{p}-value answers.
The conclusion of this section on asymmetric confidence intervals is
that a SGPV of 1 or 0 can still be interpreted as a \emph{p} \textless{}
0.025 or \emph{p} \textgreater{} 0.975 in an equivalence test, since the
SGPV and \emph{p}-value for the TOST procedure are always directly
related at the values \emph{p} = 0.025 and \emph{p} = 0.975. Although
Blume et al. (2018) state that \enquote{the degree of overlap conveys
how compatible the data are with the null premise} this definition of
what the SGPV provides does not hold for asymmetric confidence
intervals. Although a SGPV of 1 or 0 can be directly interpreted, a SGPV
between 0 and 1 is not interpretable as \enquote{compatibility with the
null hypothesis} under the assumption of normally distributed data, and
the generalizability of this statement needs to be examined beyond
normally distributed data. Indeed, Blume and colleagues write in the
supplemental material that \enquote{The magnitude of an inconclusive
second-generation \emph{p}-value can vary slightly when the effect size
scale is transformed. However definitive findings, i.e.~a \emph{p}-value
of 0 or 1 are \emph{not} affected by the scale changes.}

@@

When introducing a new statistical method, it is important to compare it
to existing approaches and specify its relative strengths and
weaknesses. First of all, even though a SGPV of 1 or 0 has a clear
interpretation (we can reject effects outside or inside the equivalence
range), intermediate values are not as easy to interpret (especially for
effects that have asymmetric confidence intervals). In one sense, they
are what they are (the proportion of overlap), but it can be unclear
what this number tells us about the data we have collected. This is not
too problematic, since the main use of the SGPV (e.g., in all examples
provided by Blume and colleagues) seems to be to examine whether the
SGPV is 0, 1, or inconclusive. As already mentioned, this interpretation
of a SGPV as is very similar to the Neyman-Pearson interpretation of an
equivalence test and a minimum effect tests (which are complementary).
The difference is that where a SGPV of 1 can be interpreted as \emph{p}
\textless{} .025, equivalence tests provide exact \emph{p}-values, and
they continue to differentiate between for example \emph{p} = 0.048 and
\emph{p} = 0.002. Whether this is desirable depends on the perspective
that is used. From a Neyman-Pearson perspective on statistical
inferences the main conclusion is based on whether or not
\(p < \alpha\), and thus an equivalence test and SGPV can be performed
by simply checking whether the confidence interval falls within the
equivalence range, just as a null-hypothesis test can be performed by
checking whether the confidence interval contains zero or not. At the
same time, it is recommended to report exact \emph{p}-values (American
Psychological Association, 2010), and exact \emph{p}-values might
provide information of interest to readers about how precisely how
surprising the data, or more extreme data, is under the null model. Some
researchers might be interested in combining an equivalence tests with a
null-hypothesis significance test. This allows a researcher to ask
whether there is an effect that is statistically different from zero,
and whether effect sizes that are considered meaningful can be rejected.
Equivalence tests combined with null-hypothesis tests classify results
into four possible categories, and for example allow researchers to
conclude an effect is significant \emph{and} equivalent (i.e.,
statistically different from zero, but also too small to be considered
meaningful, see (Lakens et al., 2018)).

@@

Blume et al. (2018) claim that \enquote{Adjustments for multiple
comparisons are obviated} (p.~15). However, this is not correct. Given
the direct relationship between TOST and SGPV highlighted in this
manuscript (where to TOST \emph{p} = 0.025 equals SGPV = 1, as long as
the SGPV is calculated based on confidence intervals, and assuming
continuous and normally distributed data), not correcting for multiple
comparisons will inflate the probability of concluding the absence of a
meaningful effect based on the SGPV in exactly the same way as for
equivalence tests. Whenever statistical tests are interpreted as support
for a hypothesis (e.g., SPGV = 0 or SGPV =1), it is possible do so
erroneously, and if researchers want to control error rates, they need
to correct for multiple comparisons.

>>>>

\documentclass[,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Equivalence Testing and the Second Generation P-Value},
            pdfauthor={Daniël Lakens~\& Marie Delacre},
            pdfkeywords={equivalence testing, second generation p-values, hypothesis testing,
TOST, statistical inference},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\shorttitle{TOST vs. SGPV}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Eindhoven University of Technology, Eindhoven, The Netherlands\\\textsuperscript{2} Service of Analysis of the Data, Université Libre de Bruxelles, Belgium}
\keywords{equivalence testing, second generation p-values, hypothesis testing, TOST, statistical inference}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\abstract{
To move beyond the limitations of null-hypothesis tests, statistical
approaches have been developed where the observed data are compared
against a range of values that are equivalent to the absence of a
meaningful effect. Specifying a range of values around zero allows
researchers to statistically reject the presence of effects large enough
to matter, and prevents practically insignificant effects from being
interpreted as a statistically significant difference. We compare the
behavior of the recently proposed second generation \emph{p}-value
(Blume, D'Agostino McGowan, Dupont, \& Greevy, 2018) with the more
established Two One-Sided Tests (TOST) equivalence testing procedure
(Schuirmann, 1987). We show that the two approaches yield almost
identical results under optimal conditions. Under suboptimal conditions
(e.g., when the confidence interval is wider than the equivalence range,
or when confidence intervals are asymmetric) the second generation
\emph{p}-value becomes difficult to interpret. The second generation
\emph{p}-value is interpretable in a dichotomous manner (i.e., when the
SGPV equals 0 or 1 because the confidence intervals lies completely
within or outside of the equivalence range), but this dichotomous
interpretation does not require calculations. We conclude that
equivalence tests yield more consistent \emph{p}-values, distinguish
between datasets that yield the same second generation \emph{p}-value,
and allow for easier control of Type I and Type II error rates.

@@

To test predictions researchers predominantly rely on null-hypothesis
tests. This statistical approach can be used to examine whether observed
data are sufficiently surprising under the null hypothesis to reject an
effect that equals exactly zero. Null-hypothesis tests have an important
limitation, in that this procedure can only reject the hypothesis that
there is no effect, while scientists should also be able to provide
statistical support for \emph{equivalence}. When testing for equivalence
researchers aim to examine whether an observed effect is too small to be
considered meaningful, and therefore is practically equivalent to zero.
By specifying a range around the null hypothesis of values that are
deemed practically equivalent to the absence of an effect (i.e., 0 ±
0.3) the observed data can be compared against an \emph{equivalence
range} and researchers can test if a meaningful effect is absent (Hauck
\& Anderson, 1984; Kruschke, 2018; Rogers, Howard, \& Vessey, 1993;
Serlin \& Lapsley, 1985; Spiegelhalter, Freedman, \& Parmar, 1994;
Wellek, 2010; Westlake, 1972).

Second generation \emph{p}-values (SGPV) were recently proposed to as a
statistic that represents \enquote{the proportion of data-supported
hypotheses that are also null hypotheses} (Blume et al., 2018). The
researcher specifies an equivalence range around a null hypothesis of
values that are considered practically equivalent to the null
hypothesis. The SGPV measures the degree to which a set of
data-supported parameter values falls within the interval null
hypothesis. If the estimation interval falls completely within the
equivalence range, the SGPV is 1. If the confidence interval falls
completely outside of the equivalence range, the SGPV is 0. Otherwise
the SGPV is a value between 0 and 1 that expresses the overlap of
data-supported hypotheses and the equivalence range. When calculating
the SGPV the set of data-supported parameter values can be represented
by a confidence interval (CI), although one could also choose to use
credible intervals or Likelihood support intervals (SI). When a
confidence interval is used, the SGPV and equivalence tests such as the
Two One-Sided Tests (TOST) procedure (Lakens, 2017; Meyners, 2012;
Quertemont, 2011; Schuirmann, 1987) appear to have close ties, because
both tests compare a confidence interval against an equivalence range.
Here, we aim to examine the similarities and differences between the
TOST procedure and the SGPV. We limit our analysis to continuous data
sampled from a bivariate normal distribution.

The TOST procedure also relies on the confidence interval around the
effect. In the TOST procedure the data are tested against the lower
equivalence bound in the first one-sided test, and against the upper
equivalence bound in the second one-sided test (Lakens, Scheel, \&
Isager, 2018). For an excellent discussion of the strengths and
weaknesses of different frequentist equivalence tests, including
alternatives to the TOST procedure, see Meyners (2012). If both tests
statistically reject an effect as extreme or more extreme than the
equivalence bound, you can conclude the observed effect is practically
equivalent to zero from a Neyman-Pearson approach to statistical
inferences. Because one-sided tests are performed, one can also conclude
equivalence by checking whether the 1-2\(\times\)\(\alpha\) confidence
interval (e.g., when the alpha level is 0.05, a 90\% CI) falls
completely within the equivalence bounds. Because both equivalence tests
as the SGPV are based on whether and how much a confidence interval
overlaps with equivalence bounds, it seems worthwhile to compare the
behavior of the newly proposed SGPV to equivalence tests to examine the
unique contribution of the SGPV to the statistical toolbox.

@@

For example, in Figure \ref{fig:TOSTSGPV4} we have plotted four SGPV's.
From A to D the SGPV is 0.76, 0.81, 0.86, and 0.91. The difference in
the percentage of overlap between A and B (-0.05) is identical to the
difference in the percentage of overlap between C and D as the mean gets
0.1 closer to the test value (-0.05). As the observed mean in a
one-sample \emph{t}-test lies closer to the test value, from situation A
to D, the mean gets closer to the test value by 0.1) the difference in
the overlap changes uniformly. As we move the observed mean closer to
the test value in steps of 0.1 across A to D the \emph{p}-value
calculated for normally distributed data are not uniformly distributed.
The probability of observing data more extreme than the upper bound of 2
is (from A to D) 0.16, 0.12, 0.08, and 0.06. As we can see, the
difference between A and B (0.04) is not the same as the difference
between C and D (0.03). Indeed, the difference in \emph{p}-values is the
largest as you start at \emph{p} = 0.5 (when the observed mean falls on
the test value), which is why the line in Figure \ref{fig:TOSTSGPV1} is
the steepest at \emph{p} = 0.5. Note that where the SGPV reaches 1 or 0,
\emph{p}-values closely approximate 0 and 1, but never reach these
values.

@@

We can examine this situation by calculating the SGPV and performing the
TOST for a situation where sample sizes are small and the equivalence
range is narrow, such that the CI is more than twice as large as the
equivalence range (see Figure \ref{fig:TOSTSGPV6}). When the two
statistics are plotted against each other we can see where the SGPV is
the same while the TOST \emph{p}-value still differentiates different
observed means (indicated by straight lines in the curve, see Figure
\ref{fig:TOSTSGPV7}). We see the SGPV is 0.5 for a range of observed
means where the \emph{p}-value from the equivalence test still varies.
It should be noted that in these calculations the \emph{p}-values for
the TOST procedure are \emph{never} smaller than 0.05 (i.e., they do not
get below 0.05 on the y-axis). In other words, we cannot conclude
equivalence based on any of the observed means. This happens because the
TOST \emph{p}-value is smaller than 0.05 only when the 90\% CI falls
completely between the upper and lower equivalence bounds. However, we
are examining a scenario where the 90\% CI is so wide that it never
falls completely within the two equivalence bounds, and thus the
equivalence test is never significant. As Lakens (2017) notes:
\enquote{in small samples (where CIs are wide), a study might have no
statistical power (i.e., the CI will always be so wide that it is
necessarily wider than the equivalence bounds).} None of the
\emph{p}-values based on the TOST procedure are below 0.05, and thus, in
the long run we have 0\% power.

@@

The \emph{p}-value from the TOST procedure still differentiates observed
means, while the SGPV does not, when the CI is wider than the
equivalence range (so the precision is low) and overlaps with the upper
and lower equivalence bound, but the CI is \emph{not} twice as wide as
the equivalence range. In the example below, we see that the CI is only
1.79 times as wide as the equivalence bounds, but the CI overlaps with
the lower and upper equivalence bounds (Figure \ref{fig:TOSTSGPV8}).
This means the SGPV is not set to 0.5, but it is constant across a range
of observed means, while the TOST \emph{p}-value is not constant across
this range.

@@

If we once more plot the two statistics against each other we see the
SGPV is 0.56 for a range of observed means where the \emph{p}-value from
the equivalence test still varies, as indicated by the straight section
of the line (Figure \ref{fig:TOSTSGPV10}).

@@

To conclude this section, there are situations where the \emph{p}-value
from the TOST procedure continues to differentiate, while the SGPV does
not. Therefore, interpreted as a continuous statistic, the SGPV is more
limited than the \emph{p}-value from the TOST procedure.

\subsection{The relation between equivalence tests and SGPV for
asymmetrical confidence intervals around
correlations}\label{the-relation-between-equivalence-tests-and-sgpv-for-asymmetrical-confidence-intervals-around-correlations}

So far we have only looked at the relation between equivalence tests and
the SGPV when confidence intervals are symmetric (e.g., for confidence
intervals around mean differences). For correlations, which are bound
between -1 and 1, confidence intervals are only symmetric for a
correlation of exactly 0. The confidence interval for a correlation
becomes increasingly asymmetric as the observed correlation nears -1 or
1. For example, with ten observations, an observed correlation of 0 has
a symmetric 95\% confidence interval ranging from -0.630 to 0.630, while
and observed correlation of 0.7 has an asymmetric 95\% confidence
interval ranging from 0.126 to 0.993. Note that calculating confidence
intervals for a correlation involves a Fisher's z-transformation, which
transforms values such that they are approximately normally distributed,
which allows one to compute symmetric confidence intervals, which are
then retransformed into a correlation, where the confidence intervals
are asymmetric if the correlation is not exactly zero.

@@

This example illustrates the difference between a proportion and a
probability. As long as data are sampled from a bivariate normal
distribution, there is always a 50\% probability of observing a
correlation smaller or larger than the true correlation, but the SGPV
for this situation depends on how far away the observed correlation is
from 0. The further away from 0, the larger the SGPV when the observed
mean falls on the equivalence bound. The SGPV is the proportion of
values in a 95\% confidence interval that overlap with the equivalence
range, but not the probability that these values will be observed. In
the most extreme case (i.e., a sample size of 4, and equivalence bounds
set to r = -0.99 and 0.99, with an observed correlation of 0.99) 58.10\%
of the confidence interval overlaps with the equivalence range, even
though in the long run only 50\% of the correlations observed in the
future will fall in this range. It should be noted that in larger sample
sizes the SGPV is closer to 0.5 whenever the observed correlation falls
on the equivalence bound, but this extreme example nevertheless clearly
illustrates the difference between question the SGPV answers, and the
question a \emph{p}-value answers. The conclusion of this section on
asymmetric confidence intervals is that a SGPV of 1 or 0 can still be
interpreted as a \emph{p} \textless{} 0.025 or \emph{p} \textgreater{}
0.975 in an equivalence test, since the SGPV and \emph{p}-value for the
TOST procedure are always directly related at the values \emph{p} =
0.025 and \emph{p} = 0.975. Although Blume et al. (2018) state that
\enquote{the degree of overlap conveys how compatible the data are with
the null premise} this definition of what the SGPV provides does not
hold for asymmetric confidence intervals. Although a SGPV of 1 or 0 can
be directly interpreted, a SGPV between 0 and 1 is not interpretable as
\enquote{compatibility with the null hypothesis} under the assumption of
a bivariate normal distribution, and the generalizability of this
statement needs to be examined beyond normal bivariate distributions.
Indeed, Blume and colleagues write in the supplemental material that
\enquote{The magnitude of an inconclusive second-generation
\emph{p}-value can vary slightly when the effect size scale is
transformed. However definitive findings, i.e.~a \emph{p}-value of 0 or
1 are \emph{not} affected by the scale changes.}

@@

When introducing a new statistical method, it is important to compare it
to existing approaches and specify its relative strengths and
weaknesses. First of all, even though a SGPV of 1 or 0 has a clear
interpretation (we can reject effects outside or inside the equivalence
range), intermediate values are not as easy to interpret (especially for
effects that have asymmetric confidence intervals). In one sense, they
are what they are (the proportion of overlap), but it can be unclear
what this number tells us about the data we have collected. This is not
too problematic, since the main use of the SGPV (e.g., in all examples
provided by Blume and colleagues) seems to be to examine whether the
SGPV is 0, 1, or inconclusive. As already mentioned, this interpretation
of a SGPV is very similar to the Neyman-Pearson interpretation of an
equivalence test and a minimum effect tests (which are complementary).
The difference is that where a SGPV of 1 can be interpreted as \emph{p}
\textless{} .025, equivalence tests provide exact \emph{p}-values, and
they continue to differentiate between for example \emph{p} = 0.048 and
\emph{p} = 0.002. Whether this is desirable depends on the perspective
that is used. From a Neyman-Pearson perspective on statistical
inferences the main conclusion is based on whether or not
\(p < \alpha\), and thus an equivalence test and SGPV can be performed
by simply checking whether the confidence interval falls within the
equivalence range, just as a null-hypothesis test can be performed by
checking whether the confidence interval contains zero or not. At the
same time, it is recommended to report exact \emph{p}-values (American
Psychological Association, 2010), and exact \emph{p}-values might
provide information of interest to readers about how precisely how
surprising the data, or more extreme data, is under the null model. Some
researchers might be interested in combining an equivalence tests with a
null-hypothesis significance test. This allows a researcher to ask
whether there is an effect that is statistically different from zero,
and whether effect sizes that are considered meaningful can be rejected.
Equivalence tests combined with null-hypothesis tests classify results
into four possible categories, and for example allow researchers to
conclude an effect is significant \emph{and} equivalent (i.e.,
statistically different from zero, but also too small to be considered
meaningful, see Lakens et al. (2018)).

@@

Blume et al. (2018) claim that \enquote{Adjustments for multiple
comparisons are obviated} (p.~15). However, this is not correct. Given
the direct relationship between TOST and SGPV highlighted in this
manuscript (where to TOST \emph{p} = 0.025 equals SGPV = 1, as long as
the SGPV is calculated based on confidence intervals, and assuming data
sampled from a continuous bivariate normal distribution), not correcting
for multiple comparisons will inflate the probability of concluding the
absence of a meaningful effect based on the SGPV in exactly the same way
as for equivalence tests. Whenever statistical tests are interpreted as
support for a hypothesis (e.g., SPGV = 0 or SGPV =1), it is possible do
so erroneously, and if researchers want to control error rates, they
need to correct for multiple comparisons.
